---
title: "UNIT 5"
author: "Mohd Musharraf Parvez"
date: "2025-04-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
```{r}
library(tidyverse)
library(forcats)
View(gss_cat)

my_data <- gss_cat %>%
  select(marital) %>%
  filter(marital %in% c("Married",
                         "Never married",
                         "Divorced")) %>%
  mutate(marital = fct_drop(marital))

my_table <- table(my_data)
View(my_table)


#Use Chi-square goodness of fitness of fit test

chisq.test(my_table)

my_data %>%
  table() %>%
  chisq.test()

#graphics

my_data %>%
  ggplot(aes(marital)) + geom_bar()
```

INTERPRETATION

The chi-squared test shows a highly significant difference (p < 2.2e-16) between the observed marital status distribution in the bar chart (where "Married" is much higher than "Never married" and "Divorced") and a hypothesized distribution (likely an equal distribution). This means the observed marital status frequencies are very unlikely to have occurred by chance if the hypothesized distribution were true.


```{r}
my_data <- gss_cat %>%
  select(race) %>%
  filter(race %in% c("White",
                        "Black", "Other")) %>%
  mutate(race = fct_drop(race))

my_table <- table(my_data)
View(my_table)


#chi sq test
chisq.test(my_table)

my_data %>%
  table() %>%
  chisq.test()

#graphics
my_data %>%
  ggplot(aes(race)) + geom_bar()


#INTERPRETATION
#The chi-squared test shows an extremely significant difference (p < 2.2e-16) between the observed racial distribution in the bar chart (where "White" is overwhelmingly dominant compared to "Black" and "Other") and a hypothesized distribution (likely an equal distribution or a pre-defined expected distribution). This indicates that the observed racial frequencies are highly unlikely to have occurred by chance if the hypothesized distribution were true.

```{r}
```
#INTERPRETATION
#The chi-squared test shows an extremely significant difference (p < 2.2e-16) between the observed racial distribution in the bar chart (where "White" is overwhelmingly dominant compared to "Black" and "Other") and a hypothesized distribution (likely an equal distribution or a pre-defined expected distribution). This indicates that the observed racial frequencies are highly unlikely to have occurred by chance if the hypothesized distribution were true.


```{r}
my_data <- gss_cat %>%
  select(partyid) %>%
  filter(partyid %in% c("Ind,near rep",
                        "Not str republican",
                        "Independent",
                        "Not str democrat",
                        "Strong democrat",
                        "Ind,near dem",
                        "Strong republican",
                        "Other party",
                        "No answer")) %>%
  mutate(partyid = fct_drop(partyid))

my_table <- table(my_data)
View(my_table)

#chi sq
chisq.test(my_table)

my_data %>%
  table() %>%
  chisq.test()

# graphics

my_data %>%
ggplot(aes(partyid)) + geom_bar()
```

INTERPRETATION

The chi-squared test reveals a highly significant difference (p < 2.2e-16) between the observed political party affiliation distribution in the bar chart (with varying counts across categories and "No answer" being extremely low) and a hypothesized distribution (likely an equal distribution or a pre-defined expected distribution). This indicates that the observed party affiliation frequencies are extremely unlikely to have occurred by chance if the hypothesized distribution were true, suggesting a strong deviation from the expected pattern.
```{r}
#religon

unique(gss_cat$relig)

my_data5 <- gss_cat %>%
  select(relig) %>%
  filter(relig %in% c(
    "Protestant",
    "Orthodox-christian",
    "None",
    "Christian",
    "Jewish",
    "Catholic",
    "Other",
    "Inter-nondenominational",
    "Hinduism",
    "Moslem/islam",
    "Buddhism",
    "No answer",
    "Native american",
    "Other eastern",
    "Don’t know"
  )) %>%

  mutate(relig = fct_drop(relig))

my_table <- table(my_data5)
View(my_table)

#chi sq
chisq.test(my_table)
my_data5 %>%
  table() %>%
  chisq.test()

# graphics

my_data5 %>%
ggplot(aes(relig)) + geom_bar()
```

INTERPRETATION

The chi-squared test shows an overwhelmingly significant difference (p < 2.2e-16) between the observed religious affiliation distribution in the bar chart (where "Protestant" is dominant, followed by "Catholic" and "None," with other religions having very low counts) and a hypothesized distribution (likely an equal distribution or a pre-defined expected distribution). This indicates that the observed religious affiliation frequencies are exceptionally unlikely to have occurred by chance if the hypothesized distribution were true, demonstrating a strong deviation from the expected pattern.
```{r}
```

```{r}
#denom

my_data <- gss_cat %>%
  select(denom) %>%
  filter(denom %in% c("Southern baptist",
                        "Baptist-dk which",
                        "No denomination",
                      "Not applicable",
                      "Lutheran-mo synod",
                      "Other",
                      "United methodist",
                      "Episcopal",
                      "Other lutheran",
                      "Afr meth ep zion",
                      "Am bapt ch in usa",
                      "Other methodist",
                      "Presbyterian c in us",
                      "Methodist-dk which")) %>%
  mutate(denom = fct_drop(denom))

my_table <- table(my_data)
View(my_table)

#chi sq
chisq.test(my_table)
my_data %>%
  table() %>%
  chisq.test()
#GRAPHICS FOR DENOM
my_data %>%
ggplot(aes(denom)) + geom_bar()
```

INTERPRETATION

The chi-squared test shows an extremely significant difference (p < 2.2e-16) between the observed denominational affiliation distribution in the bar chart (where "Not applicable" is overwhelmingly dominant, and other denominations show varying, much lower counts) and a hypothesized distribution (likely an equal distribution or a pre-defined expected distribution). This indicates that the observed denominational frequencies are exceptionally unlikely to have occurred by chance if the hypothesized distribution were true, highlighting a strong deviation from the expected pattern, primarily due to the "Not applicable" category.

```{r}
my_data6 <- gss_cat %>%
  select(tvhours) %>%
  filter(tvhours %in% c("12",
                        "NA",
                        "2",
                        "4",
                        "1",
                        "3",
                        "0",
                        "7",
                        "5",
                        "8",
                        "10",
                        "6",
                        "15",
                        "11",
                        "24",
                        "20",
                        "13",
                        "14",
                        "21",
                        "9",
                        "16",
                        "22",
                        "18",
                        "17",
                        "23")) %>%
  mutate(tvhours = as.factor(tvhours)) %>% # Convert to factor here
  mutate(tvhours = fct_drop(tvhours))

my_table <- table(my_data6)
View(my_table)

#chi sq
chisq.test(my_table)

my_data6 %>%
  table() %>%
  chisq.test()

#GRAPHICS FOR tvhours
my_data6 %>%
ggplot(aes(tvhours)) + geom_bar()
```

INTERPRETATION


The chi-squared test shows an overwhelmingly significant difference (p < 2.2e-16) between the observed distribution of TV viewing hours in the bar chart (heavily skewed towards lower hours, with a rapid decline and very low counts for higher hours) and a hypothesized distribution (likely an equal distribution or a pre-defined expected distribution). This indicates that the observed TV viewing hour frequencies are exceptionally unlikely to have occurred by chance if the hypothesized distribution were true, revealing a strong deviation from the expected pattern and highlighting the concentration of individuals watching less TV.

`